Realistically Interesting Domains for Probabilistic Planners

### 2014-04-08 FinalPaper/paper.tex
changed to comments

FF-Replan \cite{FFReplan}, a replanner based on determinization techniques and a classical planner, surprisingly outperformed other participating MDP solvers in the IPPC-04 and by doing so changed the course of the planning competition entirely. By relying on search and removing all probabilistic information from the problem, FF-replan functions with enormous speed and efficiency. This method has proven to outperform MDP solvers, the state-of-the-art probabilistic planners for uncertainty of that time. Since then, Little et al. has argued for \emph{probabilistically interesting} domains. By introducing finite horizon MDPs in IPPC-11, state of the art planners have been adapted to cope with large branching factors to approximate infinite planning problems. Because these "planning problems" can often be solved using a smaller horizon $h < H$ as introduced in Section \ref{sec:background}, they represent optimization problems rather than planning problems, favoring current state of the art planners over replanners. This section introduces the shift of the planners towards optimization  by discussing FF-replan and a current state of the art probabilistic planner \textsc{GOURMAND} and its predecessors.

@@

Adapting LRTDP to a finite horizon MDP \emph{M(H)} solver \emph{LRTDP-FH} requires for each trial to start at $s_0$ or $M(s_0, H)$ and run for at most $H$ time steps. All states $H$ steps away from $s_0$ would translate to goal states in a goal-oriented MDP. \\

@@

\subsection{\textsc{GLUTTON}}
The \textsc{GLUTTON} planner enhances LR$^2$TDP to cope with large branching factors by using several optimization techniques. It employs LR$^2$TDP in an off-line fashion by initializing the algorithm with the \emph{Max-Reward} heuristic that places an upper bound on the optimal value function $V^*$. It optimizes for large branching factors by performing the transition function in a subset of transition states $s'$. It distinguished between \emph{exogenous} effects caused by natural dynamics and \emph{pure} effects caused by the action itself and a mixed of both. It also utilized the saved memory space caused by mapping only the converged states in the hash-table by using the free space as cache for storing successor states computed using the transition function. 
Since \textsc{GLUTTON} subsamples the transition function, it may not know a good action to perform in states it missed due to subsampling. To compensate for these states, \textsc{GLUTTON} has to resort to a \emph{default policy}. \todo{Wel of niet de default policy uitleggen}.
%---------------------------------------------------------------------
%During IPPC-11 1 out of 5 planners used a LRTDP-based system and managed to outperform three other planners using the dominant algorithm for probabilistic planning: UCT. This caused to wonder whether LRTDP planners could outperform UCT based systems.  UCT forces an expert to guess a non-shortsighted lookahead value for which it should be able to converge on encountered states. Choosing this value can have very bad consequences and its existence causes the UCT to require some input, taking away from its autonomicity. At the IPPC-11 \textsc{Glutton} was an offline LRTDP based planner that came second. As PROST employed UCT in an online manner, \textsc{Gourmand} is an online adaptation of the \textsc{Glutton} algorithm to test whether LRTDP performs better if it is employed online.
%
%UCT picks an action based on it current quality estimate and an \emph{exploration term} which forces the planner to choose actions that have rarely been tried in the past. This makes UCT suitable for a wider scale of scenarios but also implies an unreliable termination condition.

@@

\subsection{\textsc{GOURMAND}}
The \textsc{GOURMAND} is an online adaptation of the \textsc{GLUTTON} planner and ensures a good \emph{anytime performance} using a lookahead value $L$ to approximate a near-optimal policy for $M(H)$ where $L < H$ and is dynamically chosen. At every iteration over the distance to the horizon $H$, it divides the total remaining time $\overline{T}$ equally amongst the remaining iterations and virtually allocates time: $T_t = \frac{\overline{T}}{H - t}$. During the first epoch, or iteration it estimates how long solving a state takes for different lookahead values. Using this estimate and a virtual time allocation \textsc{GOURMAND} determines the largest lookahead value $L_t$ for which it can almost certainly solve the current epoch  and whether it could solve for a larger lookahead $\widehat{L}_t$ by taking away time from the next epoch, still guaranteeing that it can solve both epochs. If so, it solves the current epoch for $\widehat{L}_t$ and otherwise $L_t$. This way \textsc{GOURMAND} never has to resort to default policies and converges relatively fast to a good policy in an online fashion. The \textsc{GOURMAND} planner has proven to outperform the current state of the art planners utilizing its robust any-time performance \cite{kolobov2012gourmand} in the following manner:

\begin{enumerate}
	\item Start \textsc{GOURMAND} at $s_0$
	\item Initialization of first epoch
	\begin{enumerate}
		\item Virtually allocate time for  $T_t = \frac{\overline{T}}{H}$
		\item Run LR$^2$TDP($M_s(H), T_t$)
		\item Adjust running average $\overline{T}$
	\end{enumerate}
	\item Virtually allocate time  $T_t = \frac{\overline{T}}{H - t}$
	\item Check if lookahead can be increased $\widehat{L}_t = L_t + 1$ else $\widehat{L}_t = L_t $
	\item Run LR$^2$TDP($M_s(\widehat{L}), \widehat{T}_t$)
	\item Adjust running average $\overline{T}$
	\item Determine action based on (near) optimal $\tau^*$
	\item Transition to next state
	\item Repeat steps 3 - 8 until time has run out
\end{enumerate}

>>>>

FF-Replan \cite{FFReplan}, a replanner based on determinization techniques and a classical planner, surprisingly outperformed other participating MDP solvers in the IPPC-04 and by doing so changed the course of the planning competition entirely. By relying on search and removing all probabilistic information from the problem, FF-replan functions with enormous speed and efficiency. This method has proven to outperform MDP solvers, the state-of-the-art probabilistic planners for uncertainty of that time. Since then, Little et al. has argued for \emph{probabilistically interesting} domains. By introducing finite horizon MDPs in IPPC-11, state of the art planners have been adapted to cope with large branching factors to approximate infinite planning problems. Because these ``planning problems'' can often be solved using a smaller horizon $h < H$ as introduced in Section \ref{sec:background}, they represent optimization problems rather than planning problems, favoring current state of the art planners over replanners. This section introduces the shift of the planners towards optimization  by discussing FF-replan and a current state of the art probabilistic planner \textsc{Gourmand} and its predecessors.

@@

Adapting LRTDP to a finite horizon MDP $M(H)$ solver \emph{LRTDP-FH} requires for each trial to start at $s_0$ or $M(s_0, H)$ and run for at most $H$ time steps. All states $H$ steps away from $s_0$ would translate to goal states in a goal-oriented MDP. \\

@@

\subsection{\textsc{Glutton}}
The \textsc{Glutton} planner enhances LR$^2$TDP to cope with large branching factors by using several optimization techniques. It employs LR$^2$TDP in an off-line fashion by initializing the algorithm with the \emph{Max-Reward} heuristic that places an upper bound on the optimal value function $V^*$. It optimizes for large branching factors by performing the transition function in a subset of transition states $s'$. It distinguished between \emph{exogenous} effects caused by natural dynamics and \emph{pure} effects caused by the action itself and a mixed of both. It also utilized the saved memory space caused by mapping only the converged states in the hash-table by using the free space as cache for storing successor states computed using the transition function. 
Since \textsc{Glutton} subsamples the transition function, it may not know a good action to perform in states it missed due to subsampling. To compensate for these states, \textsc{Glutton} has to resort to a \emph{default policy}. \todo{Wel of niet de default policy uitleggen}.
%---------------------------------------------------------------------
%During IPPC-11 1 out of 5 planners used a LRTDP-based system and managed to outperform three other planners using the dominant algorithm for probabilistic planning: UCT. This caused to wonder whether LRTDP planners could outperform UCT based systems.  UCT forces an expert to guess a non-shortsighted lookahead value for which it should be able to converge on encountered states. Choosing this value can have very bad consequences and its existence causes the UCT to require some input, taking away from its autonomicity. At the IPPC-11 \textsc{Glutton} was an offline LRTDP based planner that came second. As PROST employed UCT in an online manner, \textsc{Gourmand} is an online adaptation of the \textsc{Glutton} algorithm to test whether LRTDP performs better if it is employed online.
%
%UCT picks an action based on it current quality estimate and an \emph{exploration term} which forces the planner to choose actions that have rarely been tried in the past. This makes UCT suitable for a wider scale of scenarios but also implies an unreliable termination condition.

@@

\subsection{\textsc{Gourmand}}
The \textsc{Gourmand} is an online adaptation of the \textsc{Glutton} planner and ensures a good \emph{anytime performance} using a lookahead value $L$ to approximate a near-optimal policy for $M(H)$ where $L < H$ and is dynamically chosen. At every iteration over the distance to the horizon $H$, it divides the total remaining time $\overline{T}$ equally amongst the remaining iterations and virtually allocates time: $T_t = \frac{\overline{T}}{H - t}$. During the first epoch, or iteration it estimates how long solving a state takes for different lookahead values. Using this estimate and a virtual time allocation \textsc{Gourmand} determines the largest lookahead value $L_t$ for which it can almost certainly solve the current epoch  and whether it could solve for a larger lookahead $\widehat{L}_t$ by taking away time from the next epoch, still guaranteeing that it can solve both epochs. If so, it solves the current epoch for $\widehat{L}_t$ and otherwise $L_t$. This way \textsc{Gourmand} never has to resort to default policies and converges relatively fast to a good policy in an online fashion. The \textsc{Gourmand} planner has proven to outperform the current state of the art planners utilizing its robust any-time performance \cite{kolobov2012gourmand} in the following manner:

\begin{enumerate}
	\item Start \textsc{Gourmand} at $s_0$
	\item Initialization of first epoch
	\begin{enumerate}
		\item Virtually allocate time for  $T_t = \frac{\overline{T}}{H}$
		\item Run LR$^2$TDP($M_s(H), T_t$)
		\item Adjust running average $\overline{T}$
	\end{enumerate}
	\item Virtually allocate time  $T_t = \frac{\overline{T}}{H - t}$
	\item Check if lookahead can be increased $\widehat{L}_t = L_t + 1$ else $\widehat{L}_t = L_t $
	\item Run LR$^2$TDP($M_s(\widehat{L}), \widehat{T}_t$)
	\item Adjust running average $\overline{T}$
	\item Determine action based on (near) optimal $\tau^*$
	\item Transition to next state
	\item Repeat steps 3 - 8 until time has run out
\end{enumerate}
